---
title: "Copula Model Comparison: Sample Size Effects"
author: "Jesper Fischer Ehmsen^1^"
output:
  html_document:
  word_document: default
  pdf_document:
    latex_engine: xelatex
link-citations: true
linkcolor: blue
always_allow_html: true
---

# Showcase of Copula Implementation with Varying Sample Sizes

## Load packages:
```{r}
packages = c("brms","tidyverse","bayesplot",
             "pracma","here", "patchwork",
             "posterior","HDInterval",
             "loo", "furrr","cmdstanr","mnormt")


do.call(pacman::p_load, as.list(packages))

knitr::opts_chunk$set(echo = TRUE)

register_knitr_engine(override = FALSE)

set.seed(123)
```

## Overview

In this analysis, we compare three modeling approaches for bivariate normal data with correlation $\rho$:

1. **Two Marginal Gaussians** (independent, no correlation)
2. **Multivariate Gaussian** (with correlation structure)
3. **Copula Decomposition** (marginals + Gaussian copula)

**Key Feature**: All models use priors that match the simulated data:
- Means: $\mu_1 = 1$, $\mu_2 = 2$ with `normal(0, 10)` priors
- Standard deviations: $\sigma_1 = 1$, $\sigma_2 = 1$ with `normal(0, 5)` priors
- Correlation: varies across simulations

We test multiple sample sizes: **N = 50, 100, 200, 500** to show how evidence accumulates.

---

## Model Formulations

### 1. Two Marginal Gaussian Distributions (Independent)

$$
\mathbf{Y} =
\begin{bmatrix}
Y_1 \\[3pt] Y_2
\end{bmatrix}
\sim
\mathcal{N}
\!\left(
\begin{bmatrix}
\mu_1 \\[3pt] \mu_2
\end{bmatrix},
\;
\begin{pmatrix}
\sigma_1^2 & 0 \\
0 & \sigma_2^2
\end{pmatrix}
\right)
$$

### 2. Multivariate Gaussian (Correlated)

$$
\mathbf{Y} =
\begin{bmatrix}
Y_1 \\[3pt] Y_2
\end{bmatrix}
\sim
\mathcal{N}
\!\left(
\begin{bmatrix}
\mu_1 \\[3pt] \mu_2
\end{bmatrix},
\;
\begin{pmatrix}
\sigma_1^2 & \sigma_1\sigma_2\rho \\
\sigma_1\sigma_2\rho & \sigma_2^2
\end{pmatrix}
\right)
$$

When $\rho = 0$, this reduces to the independent case.

### 3. Copula Decomposition

Starting with marginals:
$$
Y_1 \sim \mathcal{N}(\mu_1,\sigma_1), \quad Y_2 \sim \mathcal{N}(\mu_2,\sigma_2)
$$

Transform to uniforms via probability integral transform, then to correlated standard normals:

$$
(z_1, z_2) = 
\begin{bmatrix}
\Phi^{-1}\big(F_1(Y_1)\big) \\
\Phi^{-1}\big(F_2(Y_2)\big)
\end{bmatrix}
\sim 
\mathcal{N}\Bigg(
\mathbf{0},
\begin{bmatrix}
1 & \rho \\
\rho & 1
\end{bmatrix}
\Bigg)
$$

---

## Stan Models

### Model 1: Two Marginal Gaussians

**Note**: Priors match simulated data parameters.

```{cmdstan, output.var="marginal"}
data {
  int<lower=1> N;             // Number of observations
  vector[N] y1;               // Outcome variable 1
  vector[N] y2;               // Outcome variable 2
}

parameters {
  vector[2] beta;                // Means [mu1, mu2]
  vector<lower=0>[2] sigma;      // Standard deviations
}

model {
  // Priors matching simulated data (mu1=1, mu2=2, sigma=1)
  beta ~ normal(0, 5);     // Weakly informative around true values
  sigma ~ normal(0, 3);     // Weakly informative around true values

  // Independent likelihoods
  for(i in 1:N){
    y1[i] ~ normal(beta[1], sigma[1]);
    y2[i] ~ normal(beta[2], sigma[2]);
  }
}

generated quantities{
  vector[N] log_lik = rep_vector(0,N);
  
  for (n in 1:N) {
    log_lik[n] += normal_lpdf(y1[n] | beta[1], sigma[1]);
    log_lik[n] += normal_lpdf(y2[n] | beta[2], sigma[2]);
  }
}
```

### Model 2: Multivariate Gaussian

**Note**: Priors match simulated data parameters.

```{cmdstan, output.var="multivariate"}
data {
  int<lower=1> N;                // Number of observations
  matrix[N, 2] y;                // Observed outcomes [y1, y2]
}

parameters {
  vector[2] beta;                // Means [mu1, mu2]
  vector<lower=0>[2] sigma;      // Standard deviations
  corr_matrix[2] Omega;          // Correlation matrix
}

transformed parameters {
  cov_matrix[2] Sigma;           // Covariance matrix
  Sigma = quad_form_diag(Omega, sigma);
}

model {
  // Priors matching simulated data (mu1=1, mu2=2, sigma=1)
  beta ~ normal(0, 5);     // Weakly informative around true values
  sigma ~ normal(0, 3);     // Weakly informative around true values
  Omega ~ lkj_corr(2);      // Uniform over correlations
  
  // Likelihood
  for (n in 1:N) {
    y[n] ~ multi_normal(beta, Sigma);
  }
}

generated quantities{
  vector[N] log_lik;
  
  for (n in 1:N) {
    log_lik[n] = multi_normal_lpdf(y[n] | beta, Sigma);
  }
}
```

### Model 3: Gaussian Copula

**Note**: Priors match simulated data parameters.

```{cmdstan, output.var="copula"}
functions {
  
  real gauss_copula_cholesky_lpdf(matrix u, matrix L) {
    array[rows(u)] row_vector[cols(u)] q;
    for (n in 1:rows(u)) {
      q[n] = inv_Phi(u[n]);
    }

    return multi_normal_cholesky_lpdf(q | rep_row_vector(0, cols(L)), L)
            - std_normal_lpdf(to_vector(to_matrix(q)));
  }
  
  vector gauss_copula_cholesky_per_row(matrix u, matrix L) {
    int N = rows(u);
    int D = cols(u);
    array[N] row_vector[D] q;
    vector[N] loglik;

    for (n in 1:N) {
        q[n,] = inv_Phi(u[n,]);
        loglik[n] = multi_normal_cholesky_lpdf(to_row_vector(q[n,]) | 
                                                 rep_row_vector(0, D), L) - 
                    std_normal_lpdf(to_vector(to_matrix(q[n,])));
    }

    return loglik;
  }
}

data {
  int<lower=0> N;
  matrix[N, 2] y;
}

parameters {
  real<lower=0> sigma1;
  real<lower=0> sigma2;
  real mu1;
  real mu2;
  
  cholesky_factor_corr[2] rho_chol;
}

transformed parameters{
  matrix[N, 2] u;
  
  for (n in 1:N) {
    u[n, 1] = normal_cdf(y[n, 1] | mu1, sigma1);
    u[n, 2] = normal_cdf(y[n, 2] | mu2, sigma2);
  }
}

model {
  // Priors matching simulated data (mu1=1, mu2=2, sigma=1)
  mu1 ~ normal(0, 5);      // Weakly informative around true values
  mu2 ~ normal(0, 5);      // Weakly informative around true values
  
  sigma1 ~ normal(0, 3);    // Weakly informative around true values
  sigma2 ~ normal(0, 3);    // Weakly informative around true values
  
  rho_chol ~ lkj_corr_cholesky(2);  // Uniform over correlations

  // Marginal likelihoods
  y[, 1] ~ normal(mu1, sigma1);
  y[, 2] ~ normal(mu2, sigma2);

  // Copula contribution
  u ~ gauss_copula_cholesky(rho_chol);
}

generated quantities {
  real rho = multiply_lower_tri_self_transpose(rho_chol)[1, 2];
  vector[N] log_lik;
  matrix[2,2] Sigma = multiply_lower_tri_self_transpose(rho_chol);
  vector[N] ll_copula = gauss_copula_cholesky_per_row(u, rho_chol);
  
  for (n in 1:N) {
    // Marginal contributions
    real ll_marg = normal_lpdf(y[n,1] | mu1, sigma1) + 
                   normal_lpdf(y[n,2] | mu2, sigma2);
    
    // Total per-trial log-likelihood
    log_lik[n] = ll_marg + ll_copula[n];
  }
}
```

---

## Simulation Study

We simulate data from a bivariate normal with:
- Means: $\mu_1 = 1$, $\mu_2 = 2$
- Standard deviations: $\sigma_1 = 1$, $\sigma_2 = 1$
- Correlations: $\rho \in [-0.6, 0.6]$ (by 0.05)
- Sample sizes: **N = 50, 100, 200, 500**

```{r, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE}

# Set up parallel plan
plan(multisession, workers = 10)

# Correlations to simulate
r = seq(-0.6, 0.6, by = 0.025)

# Sample sizes to test
sample_sizes = c(50, 100, 200, 500)

# Function to simulate and fit models
simulate_and_fit <- function(cor, n_sample) {
  
  # Simulate data with mu1=1, mu2=2, sigma1=1, sigma2=1
  n_out <- mnormt::rmnorm(
    n = n_sample, 
    mean = c(1, 2),
    varcov = cbind(c(1, cor), c(cor, 1))
  )
  
  bdata <- data.frame(
    y1 = n_out[,1],
    y2 = n_out[,2]
  )
  
  # Marginal model
  datastan_margin <- list(
    y1 = bdata$y1,
    y2 = bdata$y2,
    N = nrow(bdata)
  )
  
  fit_margin <- marginal$sample(
    data = datastan_margin,
    refresh = 0,
    iter_warmup = 2000,
    iter_sampling = 5000,
    adapt_delta = 0.99,
    parallel_chains = 4
  )
  
  # Multivariate & copula models
  datastan_multi <- list(
    y = cbind(bdata$y1, bdata$y2),
    N = nrow(bdata)
  )
  
  fit_multi <- multivariate$sample(
    data = datastan_multi,
    refresh = 0,
    iter_warmup = 2000,
    iter_sampling = 5000,
    adapt_delta = 0.99,
    parallel_chains = 4
  )
  
  fit_cor <- copula$sample(
    data = datastan_multi,
    refresh = 0,
    iter_warmup = 2000,
    iter_sampling = 5000,
    adapt_delta = 0.99,
    parallel_chains = 4
  )
  
  # Compute LOO comparison
  loo_df <- data.frame(
    loo::loo_compare(
      list(
        margin = fit_margin$loo(),
        copula = fit_cor$loo(),
        multi = fit_multi$loo()
      )
    )
  ) %>% rownames_to_column()
  
  loo_df$sim_correlation <- cor
  loo_df$sample_size <- n_sample
  
  return(list(loo_df))
}

# Create grid of parameters
param_grid <- expand.grid(
  cor = r,
  n_sample = sample_sizes
)

# Run in parallel
loos <- future_map2(
  param_grid$cor, 
  param_grid$n_sample, 
  simulate_and_fit, 
  .progress = TRUE
)

```

---

## Results

### LOO-CV Comparison Across Sample Sizes

```{r, fig.width=14, fig.height=10}
plot_data <- map_dfr(loos, 1) %>% 
  filter(elpd_diff != 0) %>% 
  mutate(models = rowname,
         sample_size = factor(sample_size, 
                             levels = c(50, 100, 200, 500),
                             labels = c("N = 50", "N = 100", "N = 200", "N = 500")))
```


```{r}
mi_theory <- expand.grid(
  r = seq(-0.6, 0.6, by = 0.01),
  sample_size = sample_sizes
) %>%
  mutate(
    mi = -0.5 * log(1 / (1 - r^2)),
    elpd_theory = mi * sample_size,
    sample_size = factor(sample_size, 
                        levels = c(50, 100, 200, 500),
                        labels = c("N = 50", "N = 100", "N = 200", "N = 500"))
  )


library(tidyverse)
library(posterior)
library(bayesplot)

font = "sans"
font_size = 16
font_size_small = 12
axis_width = 1.5
tick_width = 1.5

text = ggplot2::theme(text = ggplot2::element_text(family = font, size = font_size))
theme = theme_classic()

#patch theme
patchtheme = ggplot2::theme(plot.tag = ggplot2::element_text(size = (font_size+4),
                                                             family = "sans",
                                                             face = "bold",
                                                             hjust = 0.5,
                                                             vjust = 0.5),
                            text=ggplot2::element_text(family=font))


library(ggplot2)
library(dplyr)
library(grid)  # for unit()

# Minimal plot with coordinate system only
base_plot <- plot_data %>% 
  filter(sample_size == "N = 50" & models != "margin") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
 # geom_hline(yintercept = 0, linetype = "solid", size = 0.8, color = "black") + # x-axis
  geom_vline(xintercept = 0, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
  geom_vline(xintercept = 0.6, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
    geom_vline(xintercept = 0.4, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
    geom_vline(xintercept = 0.2, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
    geom_vline(xintercept = -0.2, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
    geom_vline(xintercept = -0.4, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
    geom_vline(xintercept = -0.6, linetype = "solid", size = 0.8, color = "black", alpha = 0) + # y-axis
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6)
  ) +
  scale_y_continuous(breaks = 0, labels = "0") +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  theme(
    axis.ticks.length = unit(5, "pt"), # optional: small ticks
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_base.tiff"),dpi = 300,width = 20, height = 14, units = "cm")



```


```{r}
plot_data %>% filter(sample_size == "N = 50" & models != "margin") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
  #  geom_line(
  #   data = mi_theory %>% filter(sample_size == "N = 50"),
  #   aes(x = r, y = elpd_theory),
  #   color = "black",
  #   linetype = "solid",
  #   alpha = 0.7,
  #   linewidth = 1.2,
  #   inherit.aes = FALSE
  # )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","orange","black"))+
    theme+text+
  theme(legend.position = "none",
        legend.key.width = unit(0.85, 'cm'), #size of bars
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))

ggsave(here::here("Mutual_info_multivar.tiff"),dpi = 300,width = 20, height = 14, units = "cm")


plot_data %>% filter(sample_size == "N = 50") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
  #  geom_line(
  #   data = mi_theory %>% filter(sample_size == "N = 50"),
  #   aes(x = r, y = elpd_theory),
  #   color = "black",
  #   linetype = "solid",
  #   alpha = 0.7,
  #   linewidth = 1.2,
  #   inherit.aes = FALSE
  # )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","black","orange"))+
    theme+text+
  theme(legend.position = "none",
        legend.key.width = unit(0.85, 'cm'), #size of bars
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_all.tiff"),dpi = 300,width = 20, height = 14, units = "cm")







plot_data %>% filter(sample_size == "N = 50") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
   geom_line(
    data = mi_theory %>% filter(sample_size == "N = 50"),
    aes(x = r, y = elpd_theory),
    color = "black",
    linetype = "solid",
    alpha = 0.7,
    linewidth = 1.2,
    inherit.aes = FALSE
  )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","black","orange"))+
    theme+text+
  theme(legend.position = "none",
        legend.key.width = unit(0.85, 'cm'), #size of bars
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_all_line.tiff"),dpi = 300,width = 20, height = 14, units = "cm")






plot_data %>% filter(sample_size == "N = 100") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
   geom_line(
    data = mi_theory %>% filter(sample_size == "N = 100"),
    aes(x = r, y = elpd_theory),
    color = "black",
    linetype = "solid",
    alpha = 0.7,
    linewidth = 1.2,
    inherit.aes = FALSE
  )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","black","orange"))+
    theme+text+
  theme(legend.position = "none",
        legend.key.width = unit(0.85, 'cm'), #size of bars
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_all_line_100.tiff"),dpi = 300,width = 20, height = 14, units = "cm")




plot_data %>% filter(sample_size == "N = 200") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
   geom_line(
    data = mi_theory %>% filter(sample_size == "N = 200"),
    aes(x = r, y = elpd_theory),
    color = "black",
    linetype = "solid",
    alpha = 0.7,
    linewidth = 1.2,
    inherit.aes = FALSE
  )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","black","orange"))+
    theme+text+
  theme(legend.position = "none",
        legend.key.width = unit(0.85, 'cm'), #size of bars
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_all_line_200.tiff"),dpi = 300,width = 20, height = 14, units = "cm")






plot_data %>% filter(sample_size == "N = 500") %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
   geom_line(
    data = mi_theory %>% filter(sample_size == "N = 500"),
    aes(x = r, y = elpd_theory),
    color = "black",
    linetype = "solid",
    alpha = 0.7,
    linewidth = 1.2,
    inherit.aes = FALSE
  )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","black","orange"))+
    theme+text+
  theme(legend.position = "none",
        legend.key.width = unit(0.85, 'cm'), #size of bars
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_all_line_500.tiff"),dpi = 300,width = 20, height = 14, units = "cm")





plot_data %>% filter(sample_size == "N = 500") %>% 
  mutate(models = ifelse(models == "margin","Independent Marginals",ifelse(models == "copula","Gaussian Copula","Multivariate Normal"))) %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff,
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, linewidth = 1.2) +
   geom_line(
    data = mi_theory %>% filter(sample_size == "N = 500"),
    aes(x = r, y = elpd_theory),
    color = "black",
    linetype = "solid",
    alpha = 0.7,
    linewidth = 1.2,
    inherit.aes = FALSE
  )+
  scale_x_continuous(
    breaks = c(-0.6,-0.4,-0.2,0,0.2,0.4,0.6), 
    labels = c("-0.6","-0.4","-0.2","0.0","0.2","0.4","0.6")
  ) +
  labs(x = "Simulated Correlation (ρ)",
    y = "ELPD Difference")+
  theme_classic() +
  scale_color_manual(values = c("red","black","orange"))+
    theme+text+
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(size=font_size_small),
        axis.text=element_text(size=font_size_small),
        axis.title=element_text(size=font_size),
        axis.line=element_line(size=axis_width),
        axis.ticks=element_line(size=axis_width))


ggsave(here::here("Mutual_info_legend.tiff"),dpi = 300,width = 20, height = 14, units = "cm")


```



```{r, fig.width=14, fig.height=10}
# Create plot
p <- plot_data %>% 
  ggplot(aes(x = sim_correlation, y = elpd_diff, color = models)) +
  geom_pointrange(aes(ymin = elpd_diff - 2*se_diff, 
                      ymax = elpd_diff + 2*se_diff),
                  size = 0.5) +
  facet_wrap(~sample_size, ncol = 2, scales = "free_y") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    text = element_text(size = 16),
    strip.text = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(
    breaks = seq(-0.6, 0.6, by = 0.2), 
    labels = seq(-0.6, 0.6, by = 0.2)
  ) +
  labs(
    x = "Simulated Correlation (ρ)",
    y = "ELPD Difference (relative to best model)",
    color = "Model",
    title = "Model Comparison via LOO-CV: Effect of Sample Size",
    subtitle = "Copula and Multivariate models are equivalent; Marginal model performs worse as |ρ| increases"
  )

print(p)
```

### Key Observations

1. **Copula ≡ Multivariate**: The copula and multivariate implementations are equivalent (randomly alternating at elpd_diff = 0)

2. **Marginal Model Degradation**: Performance loss is **quadratic** in correlation coefficient

3. **Sample Size Effects**: Larger samples provide clearer discrimination between models

---

## Mutual Information Analysis

The quadratic relationship with correlation follows from mutual information:

$$
MI(X, Y) = -\frac{1}{2} \log(1 - \rho^2)
$$

Expected ELPD difference scales with sample size and mutual information:

```{r, fig.width=14, fig.height=10}
# Add theoretical line
mi_theory <- expand.grid(
  r = seq(-0.6, 0.6, by = 0.01),
  sample_size = sample_sizes
) %>%
  mutate(
    mi = -0.5 * log(1 / (1 - r^2)),
    elpd_theory = mi * sample_size,
    sample_size = factor(sample_size, 
                        levels = c(50, 100, 200, 500),
                        labels = c("N = 50", "N = 100", "N = 200", "N = 500"))
  )

p + 
  geom_line(
    data = mi_theory,
    aes(x = r, y = elpd_theory),
    color = "black",
    linetype = "solid",
    linewidth = 1.2,
    inherit.aes = FALSE
  ) +
  labs(
    subtitle = "Black line shows theoretical MI × N prediction"
  )
```

The black line shows the theoretical expected difference based on:
$$
\text{ELPD difference} \approx -\frac{N}{2} \log(1 - \rho^2)
$$

---

## Conclusion

**Main Findings**:

1. ✅ **Copula implementation is correct**: Equivalent to multivariate normal for Gaussian marginals

2. ✅ **LOO-CV captures true model structure**: Discriminates based on dependence strength

3. ✅ **Sample size matters**: Larger N → clearer evidence for model with correct dependence structure

4. ✅ **Theoretical alignment**: Empirical results match mutual information predictions

**Next Steps**: Demonstrate copula's power with non-Gaussian marginals where multivariate normal is no longer appropriate!

